A Hedonic Price Analysis of Traffic Noise in the Twin Cities Housing Market

### 2013-06-27 reports/Draft.Rnw

\section{Data}
The 2010 US Census lists the population of the Minneapolis and St. Paul Metropolitan area in Minnesota as almost 3 million residents spread over seven counties. This study examines single family residential home transactions in the Census-defined urban areas of three of the seven counties – Dakota, Ramsey and Washington County (see Figure 1). We obtained data from approximately forty thousand sales transactions between 2005 and 2010 (n=42,095) from the 2010 MetroGIS Regional Parcel Dataset published by the Metropolitan Council. In addition to the geographic location and date of the house sale, we collected structural and locational variables commonly used in the hedonic literature.

\begin{figure}
\includegraphics[width = \textwidth]{../graphs/StudyArea_SaleValueDistribution}
\caption{This figure shows the spatial extent of our study area as well as the significant spatial variation in single family house sales prices.}
\end{figure}

@@

\begin{table}[ht]
\caption{Hedonic Regression Results (dependent variable = $ln(\textrm{Sale Price})$)} \label{tab:mainresults} 
\makebox[\linewidth]{
\small
\begin{tabular}{lrrrrrr} \\
 & \multicolumn{2}{c}{Global Model} & \multicolumn{4}{c}{Locally Weighted Regression Model} \\
 & \multicolumn{1}{c}{$\hat{\beta}$} & \multicolumn{1}{c}{$\hat{\sigma}_{\hat{\beta}}$} & mean $\hat{\beta}$ & \multicolumn{1}{c}{$\sigma_{\hat{\beta}}$} & (5) & (6) \\ 
  \hline
(Intercept) & 8.4E+00 & 1.3E-01 & 2.9E+00 & 1.5E+01 & 0.64 & 0.00 \\ 
  Traffic Noise & -2.6E-03 & 1.6E-04 & -2.5E-03 & 3.0E-03 & 0.47 & 0.67 \\ 
  House Size & 3.0E-04 & 2.7E-06 & 2.6E-04 & 9.1E-05 & 1.00 & 0.00 \\ 
  Lot Size & 2.6E-01 & 1.5E-02 & 4.0E-01 & 3.7E-01 & 0.67 & 0.00 \\ 
  Year of Construction & 1.2E-03 & 6.5E-05 & 4.5E-03 & 5.4E-03 & 0.79 & 0.00 \\ 
  Owner Occupancy & 2.7E-02 & 3.2E-03 & 2.3E-02 & 6.4E-02 & 0.37 & 1.00 \\ 
  Elementary Test Scores & 1.5E-03 & 1.2E-04 & 2.4E-04 & 3.1E-02 & 0.30 & 0.00 \\ 
  Median Income & 3.3E-06 & 7.7E-08 & 9.6E-07 & 8.1E-06 & 0.35 & 0.00 \\ 
  distance to CBD & 1.3E-05 & 6.3E-07 & 1.1E-05 & 1.0E-04 & 0.30 & 0.00 \\ 
  `` nearest Park & -3.7E-07 & 1.1E-06 & -1.6E-05 & 1.3E-04 & 0.32 & 0.00 \\ 
  `` nearest Lake & 2.5E-05 & 2.1E-06 & -2.8E-05 & 1.4E-04 & 0.35 & 0.00 \\ 
  `` nearest Shopping Center & -5.3E-06 & 1.3E-06 & 1.5E-05 & 8.8E-05 & 0.29 & 0.00 \\  
  location of sale & \multicolumn{2}{c}{city fixed effects} & \multicolumn{4}{c}{\footnotesize city fixed effects and nearest 200 sales} \\
  timing of sale & \multicolumn{2}{c}{year fixed effects} & \multicolumn{4}{c}{within last 12 months} \\
 \hline
  GCV score & \multicolumn{2}{c}{0.041} & \multicolumn{2}{c}{0.027}  &  & \\
  Moran's I statistic & \multicolumn{2}{c}{0.199} & \multicolumn{2}{c}{0.012} & & \\
\end{tabular}
}
 \caption*{\small (5) displays the proportion of $\hat{\beta}_{LWR}$ that can reject the null hypthesis of $\beta =0$.\\ (6) displays the proportion of Monte Carlo simulations with larger standard deviations of the LWR coefficients than $\sigma_{\hat{\beta}}$}
\end{table}

>>>>

\section{Data}
The 2010 US Census lists the population of the Minneapolis and St. Paul Metropolitan area in Minnesota as almost 3 million residents spread over seven counties. This study examines single family residential home transactions in the Census-defined urban areas of three of the seven counties – Dakota, Ramsey and Washington County (see Figure 1). We obtained data from approximately forty thousand sales transactions between 2005 and 2010 (n=42,095) from the 2010 MetroGIS Regional Parcel Dataset published by the Metropolitan Council. In addition to the geographic location and date of the house sale, we collected structural and locational variables commonly used in the hedonic literature. Table \ref{tab:sumStats} provides a brief description of the variables and some basic summary statistics. Figure \ref{fig:overview} shows an overview of the study area as well as the spatial distribution of the house sales prices. Table \ref{tab:cor} shows a simple correlation matrix of the quantitative variables.

\begin{figure}
\includegraphics[width = \textwidth]{../graphs/StudyArea_SaleValueDistribution}
\caption{This figure shows the spatial extent of our study area as well as the significant spatial variation in single family house sales prices.}\label{fig:overview}
\end{figure}

\begin{table}[htb]
\caption{Variable Description and Summary Statistics}\label{tab:sumStats}
\makebox[\linewidth][c]{
\small
\begin{tabular}{lrrrrrrr}

 & min & 25\% & 50\% & mean & 75\% & max & $\sigma$ \\ 
  \hline
  Sale Price (thousands \$) & 98.8 & 195.0 & 241.0 & 265.5 & 314.9 & 675.0 & 102.8 \\ 
  House Size (thousands square feet) & 390 & 1158 & 1628 & 1750 & 2188 & 4000 & 704 \\ 
  Lot Size (acres) & 0.02 & 0.17 & 0.25 & 0.25 & 0.31 & 0.60 & 0.11 \\ 
  Owner Occupancy Dummy & 0.0 & 1.0 & 1.0 & 0.8 & 1.0 & 1.0 & 0.4 \\ 
  Year House was Built & 1850 & 1950 & 1973 & 1967 & 1993 & 2010 & 32 \\ 
  Traffic Noise (dB) & 25.1 & 50.5 & 55.3 & 56.4 & 62.2 & 91.5 & 8.8 \\ 
  Median Income in Census Tract (thousands \$) & 13.9 & 54.6 & 69.1 & 72.6 & 89.7 & 143.3 & 23.6 \\ 
  Elementary Test Scores & 335.6 & 359.8 & 365.4 & 364.5 & 370.4 & 551.7 & 10.6 \\ 
  Distance to Central Business District (km) & 1.1 & 6.8 & 13.2 & 14.7 & 21.8 & 37.1 & 8.9 \\ 
  Distance to nearest Park (km) & 0.0 & 1.1 & 2.2 & 2.6 & 3.8 & 9.7 & 1.9 \\ 
  Distance to nearest Lake (km) & 0.0 & 0.4 & 0.8 & 0.9 & 1.3 & 4.4 & 0.7 \\ 
  Distance to nearest Shopping Center (km) & 0.0 & 0.9 & 1.5 & 1.9 & 2.3 & 10.8 & 1.6 \\ 
   \hline
\end{tabular}
}
\end{table}

@@

\begin{table}[bp]
\caption{Hedonic Regression Results (dependent variable = $ln(\textrm{Sale Price})$)} \label{tab:mainresults} 
\makebox[\linewidth]{
\small
\begin{tabular}{lrrrrrr} \\
 & \multicolumn{2}{c}{Global Model} & \multicolumn{4}{c}{Locally Weighted Regression Model} \\
 & \multicolumn{1}{c}{$\hat{\beta}$} & \multicolumn{1}{c}{$\hat{\sigma}_{\hat{\beta}}$} & mean $\hat{\beta}$ & \multicolumn{1}{c}{$\sigma_{\hat{\beta}}$} & (5) & (6) \\ 
  \hline
(Intercept) & 8.4E+00 & 1.3E-01 & 2.9E+00 & 1.5E+01 & 0.64 & 0.00 \\ 
  Traffic Noise & -2.6E-03 & 1.6E-04 & -2.5E-03 & 3.0E-03 & 0.47 & 0.67 \\ 
  House Size & 3.0E-04 & 2.7E-06 & 2.6E-04 & 9.1E-05 & 1.00 & 0.00 \\ 
  Lot Size & 2.6E-01 & 1.5E-02 & 4.0E-01 & 3.7E-01 & 0.67 & 0.00 \\ 
  Year of Construction & 1.2E-03 & 6.5E-05 & 4.5E-03 & 5.4E-03 & 0.79 & 0.00 \\ 
  Owner Occupancy & 2.7E-02 & 3.2E-03 & 2.3E-02 & 6.4E-02 & 0.37 & 1.00 \\ 
  Elementary Test Scores & 1.5E-03 & 1.2E-04 & 2.4E-04 & 3.1E-02 & 0.30 & 0.00 \\ 
  Median Income & 3.3E-06 & 7.7E-08 & 9.6E-07 & 8.1E-06 & 0.35 & 0.00 \\ 
  distance to CBD & 1.3E-05 & 6.3E-07 & 1.1E-05 & 1.0E-04 & 0.30 & 0.00 \\ 
  `` nearest Park & -3.7E-07 & 1.1E-06 & -1.6E-05 & 1.3E-04 & 0.32 & 0.00 \\ 
  `` nearest Lake & 2.5E-05 & 2.1E-06 & -2.8E-05 & 1.4E-04 & 0.35 & 0.00 \\ 
  `` nearest Shopping Center & -5.3E-06 & 1.3E-06 & 1.5E-05 & 8.8E-05 & 0.29 & 0.00 \\  
  location of sale & \multicolumn{2}{c}{city fixed effects} & \multicolumn{4}{c}{\footnotesize city fixed effects and nearest 200 sales} \\
  timing of sale & \multicolumn{2}{c}{year fixed effects} & \multicolumn{4}{c}{within last 12 months} \\
 \hline
  GCV score & \multicolumn{2}{c}{0.041} & \multicolumn{2}{c}{0.027}  &  & \\
  Moran's I statistic & \multicolumn{2}{c}{0.199} & \multicolumn{2}{c}{0.012} & & \\
\end{tabular}
}
 \caption*{\small (5) displays the proportion of $\hat{\beta}_{LWR}$ that can reject the null hypthesis of $\beta =0$.\\ (6) displays the proportion of Monte Carlo simulations with larger standard deviations of the LWR coefficients than $\sigma_{\hat{\beta}}$}
\end{table}
